SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-06 docs/questioning.tex
changes and questions

In this section we categorize and describe the various mechanisms
developed to detect the attacks described in
Section~\ref{sec:attacks}. We accompany our descriptions
with comments related to the following dimensions:
a) diagnostic performance, b) computational performance,
c) security, d) ease of use, and e) detection locus.
Regarding the availability dimension we provide a general
comment in our upcoming ``Analysis" section (\ref{sec:analysis}).

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches, namely:
{\it design-based}, {\it behavior-based} and {\it hybrid}.
The design-based category involves mechanisms that make applications
immune to different attack vectors by design~\cite{JL75,L81}. 
The behavior-based category incorporates a variety of
schemes that inspect the behavior of applications
and detect attacks based on specific criteria~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both categories.
Table~\ref{tab:comp} groups the various subcategories
and for every mechanism provides the following:

\begin{enumerate}
\item The number of citations of the publication(s) where
the mechanism was presented.
\item Quadruples that show the values of true
positives ({\sc tp}), true negatives ({\sc tn}),
false positives ({\sc fp}), and false negatives
({\sc tn}). For every quadruple there is a corresponding
suffix that indicates whether the testbed was based
on: real-world applications known to be vulnerable
(r), synthetic benchmarks (s), both (a).
If no testing has taken place we add a question
mark ({\bf ?}).
\item The computational overhead of the
mechanism as stated in the publication.
\item The attacks that the mechanism detects.
\end{enumerate}

@@

Table~\ref{tab:avail} presents
our findings regarding the availability of the mechanism
in terms of source code and corresponding executables.
We also examined the availability of the testbeds mentioned in the paper.
The check mark (\tick) indicates that at some point within the publication,
there is a homepage where the reader can refer to, to download the
corresponding software. {\bf AO} (Available On-line) suggests that
the software is available on-line but the address was not mentioned
in the paper, which probably indicates that the authors made
it available after the publication. The question mark ({\bf ?}),
indicates that the a homepage for the software was included
in the publication but now it is not available.

@@

Here we analyze the three basic design-based approaches
used to protect web applications from the attacks we
discussed in Section~\ref{sec:attacks}.

@@

The key idea behind this approach, is to compare
the tree representation of the abstract syntactic
structure of the source code that is about to be
executed with the one that was originally intended.
If the trees diverge, the application is probably
under attack.

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (locus {\sc z}).
Contrary to {\sc sqlg}uard (and other
mechanisms as we will observe), {\sc sql}check has been
extensively tested in terms of diagnostic performance
as it can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

@@

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that utilize
this approach to detect JavaScript-driven {\sc xss} attacks
as we will see in Subsection~\ref{sec:hybrid}.

@@

This approach is used to detect attacks that target web users.
When using a framework that implements
this approach, developers must define
specific security policies on the server side.
Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts
server responses.

{\it Noxes}~\cite{KKVJ06,KJKV09} is the only
framework that partially allows the web
user to specify policies to prevent {\sc xss} attacks.
The key idea behind Noxes is to parse
the {\sc html} that reaches the browser to
find static {\sc url} references. Then, based on
a set of policies Noxes allows or blocks
the request (locus {\sc x}). Such policies
can be also provided by the server (i.e. ``never
follow a link that leads to the malicious.com
web site"). The main issue with Noxes is that
it does not deal with JavaScript-driven {\sc xss}
attacks which is currently the most popular
{\sc xss} attack vector.

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (locus {\sc x}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
the above mechanisms is that they are vulnerable
to non-persistent {\sc xss} attacks that alter
the {\sc dom} of an already rendered
page~\cite{APKLM10}.
In addition, a recent variation of JavaScript
injection attacks known as return-to-JavaScript
attack~\cite{APKLM10} can be used to bypass
the above mechanisms. Such attacks share
many similarities with the {\it return-to-lib}
attack~\cite{SPPGMMB04}. When performing an
attack like this the malicious user transfers the
execution to points that contain script
code which in turn is used to perform the attack.

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (locus {\sc y}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protect the web user from exploitations
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (locus {\sc x}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.

@@

\noindent
{\it WebJail}~\cite{VDDPJ11}, and
{\sc soma}~\cite{OWVS08} are two frameworks that
can actually detect such an attack (locus {\sc x}).
To achieve this, {\sc soma} requires site administrators
to specify legitimate, external domains for sending
or receiving information in order to approve interactions
between them and the protected web site. As a result,
{\sc soma} can also detect {\sc csrf} attacks.
WebJail contains the functionality of third-party scripts
by introducing a web component integrator that restricts
the access that these scripts may have to either the data
or the functionality of other components.

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (locus {\sc y})
interposing on the client-server communication, and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11}
has a very similar functionality and meanwhile
addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (locus {\sc x}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

Most of the aforementioned frameworks, involve
several deployment issues since they
require significant source modifications by the
developers on the server-side, in order to
introduce the policies.

@@

{\sc isr} is a method that has been applied to counter
different kinds of application attacks~\cite{K09b,KKP03}.
The main idea behind it, is to separate code from
data by randomizing the legitimate code's 
execution environment. In this way, the malicious code
injected by the attackers who do not know the randomization
algorithm, will not get executed.

{\it {\sc sql}rand}~\cite{BK04} implements {\sc isr}
to detect {\sc sql} injection attacks. It works in the following manner:
initially, it allows programmers to create {\sc sql} statements
using randomized instructions instead of standard keywords.
The modified queries are reconstructed at runtime using
the same key that is inaccessible to the malicious user.
{\sc sql}rand is one of the few mechanisms that
detect {\sc sql} injection attacks on the database
level (locus $\Omega$).

In the case of {\sc xss} attacks that are based on JavaScript
or {\sc html}, consider a {\sc xor} function that encodes all source of a web
page on the server-side and then, on the client-side, the web browser decodes the
source by applying the same function again (locus {\sc x}).
Variations of this approach include:
{\it Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}
which randomize the instruction set of {\sc html} and
JavaScript respectively.
Contrary to x{\sc js}, in Noncespaces administrators must set
specific policies in a manner similar to a firewall
configuration language. This fact, together with the
insufficient testing of the mechanism (see Table~\ref{tab:comp})
brings up questions concerning its diagnostic performance.
{\it {\sc sm}ask}~\cite{JB07} is another framework
that was inspired by {\sc isr}. To detect {\sc xss}
attacks it searches for {\sc html} and JavaScript
keywords within the application's legitimate code.
This is done before the processing of an {\sc http}
request. When it finds one it adds a token to it.
This results to a ``code mask". Then,
before sending the resulting {\sc html} data to the
web user, the framework searches the data for
illegal code by using the same keywords (locus {\sc y}).
Since all legitimate code has been ``masked",
the injected code can be identified.
The pre- and post-processing of the code though,
may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask
did not provide measurements regarding the
computational performance of the tool (see Table~\ref{tab:comp}).

@@

Even though the above implementations impose a low
computational overhead, they impose an infrastructure
overhead. In particular, in the case of {\sc sql}rand
the integration of a proxy within the database server
is required while Noncespaces and x{\sc js} require
modifications on both the server and the client.

@@

Taint tracking is provided as a feature in some
programming languages, such as Perl and Ruby.
By enabling the feature, Perl would refuse to run code
vulnerable to an {\sc sql} injection attack
(consider a tainted variable being used in a query)
and exit with an error message.

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (locus {\sc z}).
However, this is a simplistic approach that may
lead to many false positives.

The following schemes apply further checks when
they identify that tainted data have reached a sink
(locus {\sc z}).
In particular, {\it Xu et al.}~\cite{XBS06}
tracks taint information at the level of bytes in memory.
To distinguish between legitimate and malicious uses
of untrusted data that reach a sink,
they search the data for suspicious symbols by using
regular expressions. {\sc csse}~\cite{PB05} associates
tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application
and others. When tainted data reaches a sink {\sc csse} performs
syntactic checks based on the aforementioned metadata
(locus {\sc z}). {\it {\sc php} Aspis}~\cite{PMP11} works
in a similar way. To obtain metadata, it takes advantage
of specific features provided by {\sc php}.
% Note that the above schemes detect potential attacks
% in locus {\sc z}. This automatically means that they
% cannot detect non-persistent JavaScript-driven
% {\sc xss} attacks. Although they can be
% modified to do so.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (locus {\sc y}).
A recent study showed that there are ways to circumvent
the majority of the above schemes~\cite{NBR14}.
Furthermore, most of them are not easy to deploy since
the majority of input vectors, string operations
and output vectors of the application must be instrumented.

{\it Vogt et al.}~\cite{VFJKKV07}
have developed a tainting scheme that follows a different
approach. Contrary to the aforementioned schemes that
operate on the server side, they track sensitive
information on the client side (locus {\sc x}).
This is a form of positive taint tracking, where
taint data is considered to be legitimate.
Their scheme detects JavaScript-driven {\sc xss}
attacks by ensuring that a script can send sensitive
user data only to the site from which it came from.
This though, may lead to false alarms.
{\it Stock et al.}~\cite{SLMS14} propose a scheme
that also operates in the browser (locus {\sc x}).
The scheme focuses on the detection of
{\sc dom}-based {\sc xss} attacks.
In such an attack, the malicious payload is
executed as a result of the modification of the
{\sc dom} environment so that the client-side
code runs in an unanticipated manner even if the
{\sc http} response that triggers the injection seems legitimate.
The scheme is different from the previous one because it
marks and observes data that are considered harmful.
Specifically, it employs a taint-enhanced JavaScript
engine that tracks the flow of attacker-controlled data.
To detect potential attacks the scheme utilizes
{\sc html} and JavaScript parsers that can
identify the generation of code coming from
tainted data.

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql}
or {\sc xp}ath queries) in various forms, and detect attacks
as outliers from the set of valid code statements.
An early approach, {\it {\sc didafit}}~\cite{LLW02} detects
{\sc sql} injection (locus $\Omega$) attacks by recording all database transactions.
Subsequent refinements by Valeur et al.~\cite{VMV05}
tagged each transaction with the corresponding application
as an extension of their anomaly detection framework called
{\it libAnomaly}. {\it {\sc sd}river}~\cite{MS09,MKS09,MKLS11} is
a signature-based mechanism that prevents {\sc sql} and
{\sc xp}ath injection attacks. The signatures generated
during a training phase, are based on features that can
depend either on the code statement or on its execution
environment (i.e. the stack trace).
Then, at runtime, the mechanism checks all statements for compliance
and can block code statements containing injected elements (locus {\sc z}).
Furthermore, {\sc amnesia}~\cite{HO05,HO06,HO05b} is a tool
that also detects {\sc sql} injection attacks (locus {\sc z})
by associating a query model with the location of every
{\sc sql} statement within the application.
Then at runtime, monitors the application's execution to
detect when {\sc sql} statements diverge from the expected model.

In the case of {\sc xss} attacks the various countermeasures
follow a similar pattern. {\it {\sc swap}}~\cite{WPLKK09}
creates a unique identifier ({\it script {\sc id}}) for
every legitimate script on the server.
Then, a JavaScript detection component placed in a
web proxy (locus {\sc y}) searches for injected scripts
with no corresponding {\sc id} in the server's responses.
If no injected scripts are found, the proxy forwards the
request to the client. This mechanism is relatively
inflexible since it does not support dynamic scripts.
In addition it imposes a significant overhead (see Table~\ref{tab:comp}).
The authors of {\it {\sc xssds}}~\cite{JEP08}
have implemented a similar mechanism that also supports
dynamic and external scripts. Specifically,
during training phase they build a list of all benign scripts.
In the case of external scripts they keep a white list
of all the valid domain names that contain scripts
used by the application.

The training subcategory, includes some mechanisms
that can be easily circumvented. For example,
consider an application that will send the password 
for a forgetful user via email by executing
the following query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript attack
like the one mentioned in Subsubsection~\ref{sec:policy}.

>>>>

In this section we categorize and describe the various mechanisms
developed to detect the attacks described in
Section~\ref{sec:attacks}. We accompany our descriptions with comments
related to the following dimensions: a) diagnostic performance, b)
computational performance, c) security, d) ease of use, and e)
detection locus. We treat f) availability separately in the
``Analysis" section (\ref{sec:analysis}).

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it design-based}, {\it behavior-based} and {\it hybrid}.
The design-based category involves mechanisms that make applications
immune to different attack vectors by design~\cite{JL75,L81}. 
%% PL:
%% Explain what "by design" means.
The behavior-based category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
specific criteria~\cite{D76,A00}, that is, on specific behavioral
patterns. Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

\begin{enumerate}
\item The number of times the corresponding publication has been cited.
\item Quadruples that show the values of true
positives ({\sc tp}), true negatives ({\sc tn}),
false positives ({\sc fp}), and false negatives
({\sc tn}). For every quadruple there is a corresponding
suffix that indicates whether the testbed was based
on: real-world applications known to be vulnerable
(r), synthetic benchmarks (s), both (a).
If no testing has taken place we add a question
mark ({\bf ?}).
\item The computational overhead of the
mechanism as stated in the publication.
\item The attacks that the mechanism detects.
\end{enumerate}

@@

Table~\ref{tab:avail} presents our findings regarding the availability
of the mechanism in terms of source code and corresponding
executables. We also examined the availability of the testbeds
mentioned in the paper. The check mark (\tick) indicates that at some
point within the publication, there is a homepage where the reader can
refer to, to download the corresponding software. {\bf AO} (Available
On-line) suggests that the software is available on-line but the
address was not mentioned in the paper, which probably indicates that
the authors made it available after the publication. The question mark
({\bf ?}), indicates that the a homepage for the software was included
in the publication but now it is not available.
%% PL: Do we need this paragraph here? Before we say that we tread
%% availability in the upcoming ``Analysis'' section.

@@

There exist three basic design-based approaches used to protect web
applications from executable source code injection attacks.

@@

The key idea behind this approach is to compare
the tree representation of the abstract syntactic
structure of the source code that is about to be
executed with the one that was originally intended.
If the trees diverge, the application is probably
under attack.

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (Locus {\sc z}).
Contrary to {\sc sqlg}uard (and other
mechanisms as we will observe), {\sc sql}check has been
extensively tested in terms of diagnostic performance
as it can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

@@

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that utilize
this approach to detect JavaScript-driven {\sc xss} attacks
as we will see in Subsection~\ref{sec:hybrid}.
%% PL This paragraph repeats the material of the previous one, doesn't
%% it?

@@

This approach is used to detect attacks that target web users.
%% PL All of them more or less target web users, correct?
When using a framework that implements
this approach, developers must define
specific security policies on the server side.
%% PL What kind of policies are these? Syntax-based, pattern matching?
Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts
server responses.

{\it Noxes}~\cite{KKVJ06,KJKV09} is the only
framework that partially allows the web
user to specify policies to prevent {\sc xss} attacks.
The key idea behind Noxes is to parse
the {\sc html} that reaches the browser to
find static {\sc url} references. Then, based on
a set of policies Noxes allows or blocks
the request (Locus {\sc x}). Such policies
can be also provided by the server (i.e. ``never
follow a link that leads to the malicious.com
web site"). The main issue with Noxes is that
it does not deal with JavaScript-driven {\sc xss}
attacks which is currently the most popular
{\sc xss} attack vector.

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (Locus {\sc x}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
the above mechanisms is that they are vulnerable
to non-persistent {\sc xss} attacks that alter
the {\sc dom} of an already rendered
page~\cite{APKLM10}.
In addition, a recent variation of JavaScript
injection attacks known as return-to-JavaScript
attack~\cite{APKLM10} can be used to bypass
the above mechanisms. Such attacks share
many similarities with the {\it return-to-lib}
attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
When performing an attack like this the malicious user transfers the
execution to points that contain script code which in turn is used to
perform the attack.

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (Locus {\sc y}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protects the web user from exploits
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (Locus {\sc x}).
%% PL In which way is this a kind of policy enforcement?
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.

@@

\noindent
{\it WebJail}~\cite{VDDPJ11}, and
{\sc soma}~\cite{OWVS08} are two frameworks that
can actually detect such an attack (Locus {\sc x}).
To achieve this, {\sc soma} requires site administrators
to specify legitimate, external domains for sending
or receiving information in order to approve interactions
between them and the protected web site. As a result,
{\sc soma} can also detect {\sc csrf} attacks.
WebJail contains the functionality of third-party scripts
by introducing a web component integrator that restricts
the access that these scripts may have to either the data
or the functionality of other components.

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (Locus {\sc y})
interposing on the client-server communication, and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11}
has a very similar functionality and meanwhile
addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (Locus {\sc x}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

Most of the aforementioned frameworks involve
several deployment issues since they
require significant source modifications by the
developers on the server-side in order to
introduce the policies.

@@

{\sc isr} is a method that has been applied to counter different kinds
of application attacks~\cite{K09b,KKP03}. The main idea behind it is
to separate code from data by randomizing the legitimate code's
execution environment. In this way, the malicious code injected by the
attackers who do not know the randomization algorithm, will not get
executed.

{\it {\sc sql}rand}~\cite{BK04} implements {\sc isr} to detect {\sc
  sql} injection attacks. It allows programmers to create {\sc sql}
statements using randomized instructions instead of standard keywords.
The modified queries are reconstructed at runtime using the same key
that is inaccessible to the malicious user. {\sc sql}rand is one of
the few mechanisms that detect {\sc sql} injection attacks on the
database level (Locus $\Omega$).

In the case of {\sc xss} attacks that are based on JavaScript or {\sc
  html}, consider a {\sc xor} function that encodes all source of a
web page on the server-side and then, on the client-side, the web
browser decodes the source by applying the same function again (Locus
{\sc x}). Variations of this approach include {\it
  Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}, which
randomize the instruction set of {\sc html} and JavaScript
respectively. Contrary to x{\sc js}, in Noncespaces administrators
must set specific policies in a manner similar to a firewall
configuration language. This fact, together with the insufficient
testing of the mechanism (see Table~\ref{tab:comp}) brings up
questions concerning its diagnostic performance. {\it {\sc
    sm}ask}~\cite{JB07} is another framework that was inspired by {\sc
  isr}. To detect {\sc xss} attacks it searches for {\sc html} and
JavaScript keywords within the application's legitimate code. This is
done before the processing of an {\sc http} request. When it finds one
it adds a token to it. This results to a ``code mask". Then, before
sending the resulting {\sc html} data to the web user, the framework
searches the data for illegal code by using the same keywords (Locus
{\sc y}). Since all legitimate code has been ``masked", the injected
code can be identified. The pre- and post-processing of the code
though, may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask did not provide measurements
regarding the computational performance of the tool (see
Table~\ref{tab:comp}).

@@

Even though the above implementations impose a low computational
overhead, they impose an infrastructure overhead. In particular, in
{\sc sql}rand requires the integration of a proxy within the database
server, while Noncespaces and x{\sc js} require modifications on both
the server and the client.

@@

Taint tracking is provided as a feature in some programming languages,
such as Perl and Ruby. By enabling the feature, Perl would refuse to
run code vulnerable to an {\sc sql} injection attack (consider a
tainted variable being used in a query) and would exit with an error
message.

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Locus {\sc z}).
However, this is a simplistic approach that may
lead to many false positives.
%% PL better to avoid characterisations and mention only the facts
%% (that it leads to many false positives, as shown by ...

It is possible to apply further checks when it is established that
tainted data have reached a sink (Locus {\sc z}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Locus {\sc z}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of specific features provided by {\sc
  php}.
%% PL Like what?
% Note that the above schemes detect potential attacks
% in locus {\sc z}. This automatically means that they
% cannot detect non-persistent JavaScript-driven
% {\sc xss} attacks. Although they can be
% modified to do so.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (locus {\sc y}).

A recent study showed that there are ways to circumvent the majority
of the above schemes~\cite{NBR14}. Furthermore, most of them are not
easy to deploy since the majority of input vectors, string operations
and output vectors of the application must be instrumented.

{\it Vogt et al.}~\cite{VFJKKV07} have developed a tainting scheme
that follows a different approach. Contrary to the aforementioned
schemes that operate on the server side, they track sensitive
information on the client side (Locus {\sc x}). This is a form of
positive taint tracking, where taint data is considered to be
legitimate. Their scheme detects JavaScript-driven {\sc xss} attacks
by ensuring that a script can send sensitive user data only to the
site from which it came from. This though, may lead to false alarms.
%% PL You mean a lot of false alarms?
{\it Stock et al.}~\cite{SLMS14} propose a scheme that also operates
in the browser (Locus {\sc x}). The scheme focuses on the detection of
{\sc dom}-based {\sc xss} attacks. In such an attack, the malicious
payload is executed as a result of the modification of the {\sc dom}
environment so that the client-side code runs in an unanticipated
manner even if the {\sc http} response that triggers the injection
seems legitimate. The scheme is different from the previous one
because it marks and observes data that are considered harmful.
Specifically, it employs a taint-enhanced JavaScript engine that
tracks the flow of attacker-controlled data. To detect potential
attacks the scheme utilizes {\sc html} and JavaScript parsers that can
identify the generation of code coming from tainted data.

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Locus
$\Omega$) attacks by recording all database transactions. Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e. the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Locus {\sc z}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (locus {\sc z}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

Various countermeasures against {\sc xss} attacks follow a similar
pattern. {\it {\sc swap}}~\cite{WPLKK09} creates a unique identifier
({\it script {\sc id}}) for every legitimate script on the server.
Then, a JavaScript detection component placed in a web proxy (Locus
{\sc y}) searches for injected scripts with no corresponding {\sc id}
in the server's responses. If no injected scripts are found, the proxy
forwards the request to the client. This mechanism is relatively
inflexible since it does not support dynamic scripts. In addition it
imposes a significant overhead (see Table~\ref{tab:comp}). The authors
of {\it {\sc xssds}}~\cite{JEP08} have implemented a similar mechanism
that also supports dynamic and external scripts. Specifically, during
training phase they build a list of all benign scripts. In the case of
external scripts they keep a white list of all the valid domain names
that contain scripts used by the application.

Defenses based on training include some mechanisms that can be easily
circumvented. For example, consider an application that will send the
password for a forgetful user via email by executing the following
query: \bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript attack
like the one mentioned in Subsubsection~\ref{sec:policy}.
%% PL Like the return-to-lib attack?
